---
title: "ht_extract: Homogeneity of Effect"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ht_extract: Homogeneity of Effect}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r, warning = FALSE, message = FALSE}
library(mdrcAnalysis)
library(dplyr)
```

## Overview and Motivation

The HT statistic (`ht_extract`) is a test of homogeneity of impact. The HT statistic is used to check whether impacts for subgroups are statistically different.

## Parameters

The `ht_extract()` function as currently implemented only requires a data set. However, the input data set must have the `SubGrpVal` column. This column is created whenever the `.subgroup=` parameter is used in other extract functions. That is, users are expected to calculate the necessary statistics using `hetreg_extract()`, `clusterreg_extract()`, or `lm_extract()` with the appropriate `.subgroup=` specification and then pass the resulting extract/data frame to `ht_extract()`. 


## Example with BIF data
The sample BIF data is a synthetic data set with baseline characteristic data from participants. These data are not typically used in subgroup analysis, but the construction of functions is similar. 

### Prep sample data and vectors of column names
NOTE: Missing values in the COHORT variable are removed because COHORT is used as the subgroup and only one participant is missing COHORT. In cases where more participants are missing the subgroup variable, it may be possible to leave them in the analysis and have the missing level treated as a separate subgroup. However, this requires sufficient sample to have differences calculated within the missing subgroup. 

```{r}
# Convert COHORT to a factor for analysis
finalBif <- bif_final %>%
  dplyr::filter(!is.na(COHORT))

# Create vector of dependent variables
finalBif_dependents <- names(finalBif[setdiff(names(finalBif), 
                                              c("SAMPLEID", "RA_CODE", "CohortFactor", 
                                                "COHORT", "RA_DATE", "AGE", "DOB", "blcurhrs_73plus"))]) 
finalBif_treatment  <- "RA_CODE"
```

## Run extracts
As shown below, `hetreg_extract()` is run first with `.subgroup = "COHORT"`. The output from that extract is then passed to `ht_extract()`. This process is the same for output from `clusterreg_extract()` and `lm_extract()`.
```{r, results=FALSE}
ht_extract_r <-
  hetreg_extract(.dataset = finalBif
                ,.subgroup = "COHORT"
                ,.dependents = finalBif_dependents
                ,.treatment = "RA_CODE"
                ,.covariates = NA_character_
                ) %>% 
  ht_extract()
```

```{r}
ht_extract_r
```

Additionally, it is possible to save the intermediate extract and then pass the object explicitly to `ht_extract()`

```{r, results = FALSE}
first_extract <-
  hetreg_extract(.dataset = finalBif
                ,.subgroup = "COHORT"
                ,.dependents = finalBif_dependents
                ,.treatment = "RA_CODE"
                ,.covariates = NA_character_
                )

second_extract <- 
  ht_extract(.dataset = first_extract)
```

```{r}
second_extract
```

